Hyperplane Elimination for Quickly Enumerating Local Optima

### 2015-09-17 Hyperplane Enumeration/main.tex
Changed from figure* to figure
Conclusion and code URL

\section{Hyperplane Elimination}
\begin{figure*}
  \centering
  \includegraphicsfit{Enumerate}
  \caption{Example change of enumeration ordering. The gray variables represent all dependencies
           for move $m_i$ which flips variable $F$. By reordering, $m_i$'s lowest $index$ dependency improves from 2 to 4.}
  \label{fig-enumerate}
\end{figure*}

@@

Each algorithm was given a maximum of four hours on a 2.5GHz Intel Xeon E5-2670v2 processor to find
all $r$-bit local optima. Any run which failed was rerun once to prevent cluster issues from introducing
noise. For the same reason any run which was over ten times slower than the next slowest run of the
same configuration with a different seed was rerun. This resulted in 33 out of 36,000 runs being rerun.
If the algorithm failed to complete in the alloted time for an instance,
we declared it unsuccessful for that combination of $r$ and $N$ of that problem. All results
we report are for successful combinations. For timing purposes local optima were counted but not
recorded.

@@

\begin{figure*}
  \centering
  \includegraphicsfit{boxplot-method}
  \caption{Comparison of completion time variance for the largest size of each problem
           where all three methods were successful at finding all 1-bit local optima.}
  \label{fig-boxplot-method}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphicsfit{length-method}
  \caption{Comparison of how each method scales with problem size when finding 1-bit local optima
           on log-linear scales. Each point is the mean
           runtime over 30 instances and each line the linear model. All confidence intervals too tight to see.}
  \label{fig-length-method}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphicsfit{slope-method}
  \caption{Estimated slope with 95\% confidence intervals for each method to find $r$-bit local optima. Slope is $m$ in the model $y = 2^c2^{mN}$
           where $y$ is number of seconds required to
	         find all local optima of a problem using $N$ bits.}
  \label{fig-slope-method}
\end{figure*}

@@

\begin{figure*}
  \centering
  \includegraphicsfit{length-radius}
  \caption{Number of $r$-bit local optima for each size of each problem. Note that
           the number of local optima in Concatenated Traps does not change while $r<k$.}
  \label{fig-length-radius}
\end{figure*}

@@

\section{Conclusions}
TODO Consider discussing neutral networks

\subsubsection*{Acknowledgments.} TODO Acknowledge BEACON.

>>>>

\section{Hyperplane Elimination}
\begin{figure}
  \centering
  \includegraphicsfit{Enumerate}
  \caption{Example change of enumeration ordering. The gray variables represent all dependencies
           for move $m_i$ which flips variable $F$. By reordering, $m_i$'s lowest $index$ dependency improves from 2 to 4.}
  \label{fig-enumerate}
\end{figure}

@@

Each algorithm was given a maximum of four hours on a 2.5GHz Intel Xeon E5-2670v2 processor to find
all $r$-bit local optima. Any run which failed was rerun once to prevent cluster issues from introducing
noise. For the same reason any run which was over ten times slower than the next slowest run of the
same configuration with a different seed was rerun. This resulted in 33 out of 36,000 runs being rerun.
If the algorithm failed to complete in the alloted time for an instance,
we declared it unsuccessful for that combination of $r$ and $N$ of that problem. All results
we report are for successful combinations. For timing purposes local optima were counted but not
recorded. All of our code, data, and statistical analysis can be downloaded from our
website.~\footnote{\url{https://name.of.site/removed/to/be/anonymous/}}

@@

\begin{figure}
  \centering
  \includegraphicsfit{boxplot-method}
  \caption{Comparison of completion time variance for the largest size of each problem
           where all three methods were successful at finding all 1-bit local optima.}
  \label{fig-boxplot-method}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphicsfit{length-method}
  \caption{Comparison of how each method scales with problem size when finding 1-bit local optima
           on log-linear scales. Each point is the mean
           runtime over 30 instances and each line the linear model. All confidence intervals too tight to see.}
  \label{fig-length-method}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphicsfit{slope-method}
  \caption{Estimated slope with 95\% confidence intervals for each method to find $r$-bit local optima. Slope is $m$ in the model $y = 2^c2^{mN}$
           where $y$ is number of seconds required to
	         find all local optima of a problem using $N$ bits.}
  \label{fig-slope-method}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphicsfit{length-radius}
  \caption{Number of $r$-bit local optima for each size of each problem. Note that
           the number of local optima in Concatenated Traps does not change while $r<k$.}
  \label{fig-length-radius}
\end{figure}

@@

\section{Conclusions and Future Work}
By expressing a function as an Mk Landscapes, we have developed a method to find all
local optima of that function very efficiently. The structure imposed by
an Mk Landscape allows for the identification of hyperplanes in the search space
which cannot contain any local optima. Combined with a greedy method for
changing the order of how the space is enumerated, we were able to prove
this method's runtime complexity scales with the number of local optima, unlike
previous methods which scale with the number of possible solutions.
When applied to NKq Landscapes and Ising Spin Glasses this allowed for a dramatic
increase in the size of problems which can be enumerated. We believe that
these larger problem sizes will be more representative of the types
of search spaces where optimization is performed.

We show how this hyperplane elimination can be extended to find only solutions
which cannot be improved by flipping $r$ or fewer bits. For some problems
this counterintuitively reduces search times, and on Adjacent NKq allowed
us to find all 4-bit local optima of a 200 bit problem.

The obvious next step is to use the results of hyperplane elimination
in conjunction with landscape analysis
techniques~\cite{tomassini:2008:nknetworks,verel:2011:nknetworks,ochoa:2015:crossovernetworks}.
We are especially interested to see how landscape properties change for larger $N$
and $r$. As to the algorithm itself, we suspect there may be room for improvement
in enumeration ordering. While we are able to prove that all optimal orderings
are greedy, the reverse is not true due to how ties are broken. Therefore more
gains may be possible by improving this ordering. Furthermore, we assumed
that all moves are equally likely to be fitness improving. It is likely possible
to determine the true probability of a move being fitness improving, resulting
in potentially better ordering.

\subsubsection*{Acknowledgments.}
This material is based in part upon work supported by the National Science Foundation
under Cooperative Agreement No. DBI-0939454. Any opinions, findings, and conclusions
or recommendations expressed in this material are those of the author(s) and do not
necessarily reflect the views of the National Science Foundation.
